\section{Double Summations and Products of Infinite Series}

\bx{ % 2.8.1
    Using tha particular array $(a_{ij})$ from Section 2.1, compute $\lim_{n \rightarrow \infty} s_{nn}$. How does this compare to the two iterated values for the sum already computed?
    \sol{
        To recall, the particular array in Section 2.1 is \[\begin{bmatrix}
            -1 & \frac12 & \frac14 & \frac18 & \frac1{16} & \dots \\[4pt] 
            0 & -1 & \frac12 & \frac14 & \frac18 & \dots \\[4pt]
            0 & 0 & -1 & \frac12 & \frac14 & \dots \\[4pt]
            0 & 0 & 0 & -1 & \frac12 & \dots \\[4pt]
            0 & 0 & 0 & 0 & -1 & \dots \\
            \vdots & \vdots & \vdots & \vdots & \vdots & \ddots \\
        \end{bmatrix}\] Doing some calculations, $s_{nn} = -2 + 1/2^{n-1}$, so $\lim s_{nn} = -2$. This is the same as when fixing $j$ and summing down each column, since each column has finitely many non-zero elements.
    }
}

\bx{ % 2.8.2
    Show that if the iterated series \[
    \sum_{i=1}^{\infty} \sum_{j=1}^{\infty} |a_{ij}|\]converges (meaning that for each fixed $i \in \mathbb{N}$ the series $\sum_{j=1}^{\infty} |a_{ij}|$ converges to some real number $b_i$, and the series $\sum_{i=1}^{\infty} b_i$ converges as well), then the iterated series \[
    \sum_{i=1}^{\infty} \sum_{j=1}^{\infty} a_{ij}\]converges. 
    \sol{
        This is obvious. We can use triangle inequality on the countable infinite terms in the double summation. To be more rigourous, let $c_i = \sum_{j=1}^\infty a_{ij}$. Since $\sum_{j=1}^\infty |a_{ij}|$ converges, $c_o$ converges as well. With $|c_i| \leq |b_i|$, then $\sum_{i=1}^{\infty} c_i$ converges by comparison text with $\sum_{i=1}^\infty |b_i|$. 
    }
}

\bx{ % 2.8.3
    \ea{
        \item Prove that $(t_{nn})$ converges.
        \item Now, use the fact that $(t_{nn})$ is a Cauchy Sequence to argue that $(s_{nn})$ converges.
    }
    \sol{\ea{
        \item Since $|a_{ij}| > 0$, $t_{nn}$ is monotone increasing. Hence for MCT to tell us it converges, we just need to show that the sequence is bounded.\[
        \sum_{i=1}^{\infty} \sum_{j=1}^{\infty} |a_{ij}| \geq \sum_{i=1}^{\infty} \sum_{j=1}^{n} |a_{ij}| \geq \sum_{i=1}^{n} \sum_{j=1}^n |a_{ij}| = t_{nn}\]
        
        \item Since $(t_{nn})$ is a Cauchy Sequence, for any $\epsilon > 0$ there exists $N$ such that if $p>q>N$, \begin{align*}
            \epsilon > |t_{pp} - t_{qq}| & = \left|\sum_{i=1}^{p} \sum_{j=1}^{p} |a_{ij}| - \sum_{i=1}^{q} \sum_{j=1}^{q} |a_{ij}| \right| \\
            & = \left| \sum_{i=q+1}^{p} \sum_{j=1}^{p} |a_{ij}| + \sum_{i=1}^{q} \sum_{j=q+1}^{p} |a_{ij}| + \sum_{i=1}^q \sum_{j=1}^{q} |a_{ij}| - \sum_{i=1}^{q} \sum_{j=1}^q  |a_{ij}| \right| \\
            & = \left| \sum_{i=q+1}^{p} \sum_{j=1}^{p} |a_{ij}| + \sum_{i=1}^{q} \sum_{j=q+1}^{p} |a_{ij}| \right| \\
            & \geq \left| \sum_{i=q+1}^{p} \sum_{j=1}^{p} a_{ij} + \sum_{i=1}^{q} \sum_{j=q+1}^{p} a_{ij} \right| \\
            & = \left| \sum_{i=q+1}^{p} \sum_{j=1}^{p} a_{ij} + \sum_{i=1}^{q} \sum_{j=q+1}^{p} a_{ij} + \sum_{i=1}^q \sum_{j=1}^{q} a_{ij} - \sum_{i=1}^{q} \sum_{j=1}^q  a_{ij} \right| \\
            & = \left|\sum_{i=1}^{p} \sum_{j=1}^{p} a_{ij} - \sum_{i=1}^{q} \sum_{j=1}^{q} a_{ij} \right| \\
            & = |s_{pp} - s_{qq}|
        \end{align*}thus $s_{nn}$ is also a Cauchy Sequence, hence converging.
    }}
}

\bx{ % 2.8.4
    \ea{
        \item Let $\epsilon > 0$ be arbitrary and argue that there exists an $N_1 \in \mathbb{N}$ such that $m,n \geq N_1$ implies $B - \frac{\epsilon}{2} < t_{mn} \leq B$.
        \item Now, show that there exists an $N$ such that \[
        |s_{mn} - S| < \epsilon\]for all $m,n \geq N$.
    }
    \sol{\ea{
        \item With $B$ being the upper bound on $\{t_{mn} : m,n\in \mathbb{N}\}$, we get $t_{mn} \leq B$. Lemma 1.3.8 tells us that there exists some $p,q$ such that $t_{pq} > B - \epsilon/2$. And since $p_1 \leq p_2$ and $q_1 \leq q_2 \implies t_{p_1q_1} \leq t_{p_2q_2}$, we can choose $N_1 = \max \{p,q\}$.
        \item By triangle inequality, $|s_{mn} - S| \leq |s_{mn} - s{nn}| + |s_{nn} - S|$. WLOG, let $m > n$, \[
        |s_{mn} - s{nn}| = \left|\sum_{i = n}^{m} \sum_{j=1}^{n} a_{ij}\right| \leq \sum_{i=n}^{m} \sum_{j-1}^{n} |a_{ij}| = |t_{mn} - t_{nn}| \leq \frac\epsilon2 \]from part (a), as long as $m,n \geq N_1$. Since $S= \lim s_{nn}$ there exists $N_2$ such that for $n \geq N_2$, $|s_{nn} - S| < \epsilon/2$, thus picking $N = \max \{N_1, N_2\}$ satisfies that condition.
    }}
}

\bx{ % 2.8.5
    \ea{
        \item Show that for all $m \geq N$ \[
        |(r_1 + r_2 + \dots + r_m) - S| \leq \epsilon\]Conclude that the iterated sum $\sum_{i=1}^{\infty} \sum_{j=1}^{\infty} a_{ij}$ converges to $S$.
        \item Finish the proof by showing that the other iterated sum, $\sum_{j=1}^{\infty} \sum_{i=1}^{\infty} a_{ij}$, converges to $S$ as well. Notice the same argument can be used once it is established that, for each fixed column $j$, the sum $\sum_{i=1}^{\infty} a_{ij}$ converges to some real number $c_j$.
    }
    \sol{\ea{
        \item Suppose $\sum_{j=1}^{\infty} a_{ij}$ converges to some value $r_i$. We can choose $N_3$ large enough such that when $n> N_3$ and $k \leq m$, we get \[
        \left|r_k - \sum_{j=1}^{n} a_{kj}\right| \leq \frac{\epsilon}{2m}\] Then there must exists some $N > N_3$ such that when $m,n \geq N$, \begin{align*}
            \left| \sum_{i=1}^{m}r_i - S\right| & \leq \left| \sum_{i=1}^{m} - \sum_{i=1}^{m} r_i \sum_{j=1}^{n} a_{ij}\right| + \left| \sum_{i=1}^{m} \sum_{j=1}^{n} a_{ij} - S \right| \\
            & = \left| \sum_{i=1}^{m} \left( r_i - \sum_{j=1}^{n} a_{ij} \right) \right| + |s_{mn} - S| \\
            & \leq \sum_{i=1}^{m} \left| r_i - \sum_{j=1}^{n} a_{ij} \right| + | s_{mn} - S| \\
            & < \sum_{i=1}^{m} \left( \frac{\epsilon}{2m} \right) + \frac{\epsilon}{2} = \epsilon
        \end{align*} and thus $\sum_{i=1}^{\infty} \sum_{j=1}^{\infty} a_{ij}$ converges to $S$.

        \item To show that $\sum_{j=1}^{\infty} \sum_{i=1}^{\infty} a_{ij}$ converges to the same value $S$, we will use the same argument. Our hypothesis guarantees that for each fixed column $j$, the series $\sum_{i=1}^{\infty} a_{ij}$ converges absolutely to some real number $c_j$. Using the same argument as in (a), we can show that \[
        \left| (c_1 + c_2 + \dots + c_m) - S\right| \leq \epsilon\] for all $m \geq N$. Hence, the iterated sum $\sum_{j=1}^{\infty} \sum_{i=1}^{\infty} a_{ij}$ converges to the same value of $S$.
    }}
}

\bx{ % 2.8.6
    \ea{
        \item Assuming the hypothesis -- and hence the conclusion -- of Theorem 2.8.1, show that $\sum_{k=2}^{\infty} d_k$ converges absolutely.
        \item Imitate the strategy in the proof of Theorem 2.8.1 to show that $\sum_{k=2}^{\infty}$ converges to $S = \lim_{n\rightarrow \infty} s_{nn}$.
    }
    \sol{\ea{
        \item The special case of Theorem 2.8.1 is if $\sum_{i=1}^{\infty} \sum_{j=1}^{\infty} |a_{ij}|$ converges, then $\sum_{k=2}^{\infty} d_k$ converges to the same values as $\sum_{i=1}^{\infty} \sum_{j=1}^{\infty} a_{ij}$ and $\sum_{j=1}^{\infty} \sum_{i=1}^{\infty} a_{ij}$.

        It is easy to see that $\sum_{i=1}^{\infty} \sum_{j=1}^{\infty} a_{ij}$ contains all the terms of $\sum_{k=2}^{\infty} a_{ij}$. By comparison to $\sum_{i=1}^{\infty} \sum_{j=1}^{\infty} |a_{ij}|$, $\sum_{k=2}^{\infty}$ must converge.

        \item We will first note that $\sum_{k=2}^{n} d_k$ contains all the terms of $s_{pp}$ when $p \leq n/2$, and $s_{qq}$ contains all the terms of $\sum_{k=2}^{n} d_k$ when $q \geq n-1$. Intuitively, we can use Squeeze Theorem to show our desired sum converges to $S$. However, we will be following a more rigorous path to form our proof as desired by the question.
        
        We will need to show that for all $\epsilon > 0$ there exists $N>0$ such that for all $n> N$, $|S - \sum_{k=2}^{\infty} d_k | < \epsilon$. Since $(s_{nn}) \rightarrow S$, for an arbitrary $\epsilon_1 > 0$, we can choose $n_1$ large enough such that $|s_{n_1n_1} - S| < \epsilon_1$. If we choose $N > n_1$, then whenever $n > N$, $\sum_{k=2}^{n} d_k$ will contain all the terms in $s_{n_1n_1}$, and if we choose $m = n-1$, then $s_{mm}$ will contain all terms in $\sum_{k=2}^{n} d_k$. Thus we get \[
        \sum_{k=2}^{n} |d_k| - t_{n_1n_1} \leq t_{mm} - t_{n_1n_1}\] As shown in Exercise 2.8.3a, $t_{nn}$ converges and is hence a Cauchy Sequence. For an arbitrary $\epsilon_2 > 0$, we can choose $n_2$ large enough to ensure that for any $m_1 > n_1$, $t_{m_1m_1} - t_{n_1n_1} < \epsilon_2$.

        Putting all these together and choosing $\epsilon_1 = \epsilon_2 = \epsilon/2$, we can choose a $N > n_1, n_2$ to give us \begin{align*}
            \left| S - \sum_{k=2}^{n} d_k \right| & \leq |S - s_{nn}| + \left| \sum_{k=2}^{n} d_k - s_{nn} \right| \\
            & < \epsilon_1 + \left| \sum_{k=2}^{n} |d_k| - t_{nn} \right| \\
            & \leq \epsilon_1 + (t_{mm} - t_{nn}) < \epsilon_1 + \epsilon_2 \\
            & = \epsilon
        \end{align*} thus completing the proof.
    }}
}

\bx{ % 2.8.7
    Assume that $\sum_{i=1}^{\infty} a_i$ converges absolutely to $A$, and $\sum_{j=1}^{\infty} b_j$ converges absolutely to $B$. \ea{
        \item Show that the iterated sum $\sum_{i=1}^{\infty} \sum_{j=1}^{\infty} |a_ib_j|$ converges so that we may apply Theorem 2.8.1.
        \item Let $s_{nn} = \sum_{i=1}^{n} \sum_{j=1}^{n} a_ib_j$, and prove that $\lim_{n\rightarrow\infty} s_{nn} = AB$. Conclude that \[
        \sum_{i=1}^{\infty} \sum_{j=1}^{\infty} a_i b_j = \sum_{j=1}^{\infty}  \sum_{i=1}^{\infty} a_i b_j = \sum_{k=2}^{\infty} d_k = AB\] where, as before, $d_k = a_1 b_{k-1} + a_2 b_{k-2} + \dots + a_{k-1} b_1$.
    }
    \sol{\ea{
        \item Let $\sum_{i=1}^{\infty} |a_i| = A'$ and $\sum_{j=1}^{\infty} |b_j| = B'$. By Algebraic Limit Theorem for Series, \[
        \sum_{i=1}^{\infty} \sum_{j=1}^{\infty} |a_i b_j| = \sum_{i=1}^{\infty} \left(|a_i| \sum_{j=1}^{\infty} |b_j|\right) = \left(\sum_{i=1}^{\infty} |a_i| \right) \left( \sum_{j=1}^{\infty} |b_j| \right) = A'B'\]

        \item Theorem 2.8.1 tells us the different iterated sums have the same limit. Using the same manipulation as above, the limit is $AB$.
    }}
}