\section{The Cauchy Criterion}

\bx{ % 2.6.1
    Supply a proof for Theorem 2.6.2.
    \sol{
        Suppose $(x_n) \rightarrow x$, hence for $\epsilon > 0$, $|x - x_n| < \epsilon/2$ when $n > N$. Similarly, $|x_m - x| < \epsilon/2$ for $m > N$. By Triangle Inequality, \[
        |x_n - x_m| \leq |x_n - x| + |x - x_m| < \epsilon/2 + \epsilon/2 = \epsilon\]
    }
}

\bx{ % 2.6.2
    Give an example of each of the following, or argue that such a request is impossible. \ea{
        \item A Cauchy sequence that is not monotone.
        \item A Cauchy sequence with an unbounded subsequence.
        \item A divergent monotone sequence with a Cauchy subsequence.
        \item An unbounded sequence containing a subsequence that is Cauchy.
    }
    \sol{\ea{
        \item $x_n = (-1)^n/n$ converges and hence Cauchy but not monotone.
        \item Impossible. Cauchy sequences must converge and hence bounded.
        \item Impossible. A Cauchy subsequence will converge, implying that it is bounded. Since the sequence is montone, it is also bounded, thus MCT tells us it converges.
        \item $(2, 1/2, 3, 1/3, \dots)$ has the subsequence $(1/2, 1/3, \dots)$ which is Cauchy.
    }}
}

\bx{ % 2.6.3
    If \seq{x}{n} and \seq{y}{n} are Cauchy sequences, then one easy way to prove that $(x_n + y_n)$ is Cauchy is to use the Cauchy Criterion. By Theorem 2.6.4, \seq{x}{n} and \seq{y}{n} must be convergent, and the Algebraic Limit Theorem then implies $(x_n + y_n)$ is convergent and hence Cauchy. \ea{
        \item Give a direct argument that $(x_n + y_n)$ is a Cauchy sequence that does not use the Cauchy Criterion or the Algebraic Limit Theorem.
        \item Do the same for the product $(x_ny_n)$.
    }
    \sol{\ea{
        \item $|(x_n + y_n) - (x_m + y_m)| = |(x_n - x_m) + (y_n - y_m)| \leq |x_n - x_m| + |y_n -y_m| < \epsilon /2 + \epsilon /2 = \epsilon$.
        \item Let $|x_n| \leq M_x$ and $|y_n| \leq M_y$ for all $n \in \mathbb{N}$. \begin{align*}
            |x_ny_n - x_my_m| &= |x_ny_n - x_ny_m + x_ny_m - x_my_m| \\
            &\leq |x_n (y_n - y_m)| + |y_m (x_n - x_m)|= |x_n||y_n - y_m| + |y_m||x_n-x_m| \\
            &\leq M_x |y_n - y_m| + M_y |x_n - x_m| = M_x (\frac{\epsilon}{2M_x}) + M_y (\frac{\epsilon}{2M_y}) = \epsilon
        \end{align*}After setting $|x_n -x_m| < \epsilon / (2M_y)$ and $|y_n -y_m| < \epsilon/(2M_x)$.
    }}
}

\bx{ % 2.6.4
    Let \seq{a}{n} and \seq{b}{n} be Cauchy sequences. Decide whether each of the following sequences is a Cauchy sequence, justifying each conclusion. \ea{
        \item $c_n = |a_n - b_n|$
        \item $c_n = (-1)^n a_n$
        \item $c_n = [[a_n]]$, where $[[x]]$ refers to the greatest integer less than or equal to $x$.
    }
    \sol{\ea{
        \item By Triangle Inequality and that $||x| - |y|| \leq |x-y|$ as proven in earlier Exercises, \begin{align*}
            |c_n - c_m| & = ||a_n - b_n| - |a_m - b_m|| \\
            &\leq |a_n - b_n -a_m + b_m| = |a_n - a_m + b_m - b_n| \\
            &\leq |a_n - a_m| + |b_m -b_n| < \epsilon / 2 + \epsilon / 2 = \epsilon
        \end{align*}
        \item No. If $a_n = 1$, then $(-1)^n$ diverges and thus not Cauchy.
        \item No. If $a_n = 1 - (-1)^n/n$, then $c_n$ oscillate between 0 and 1, thus not Cauchy.
    }}
}

\bx{ % 2.6.5
    Consider the following (invented) definition: A sequence $(s_n)$ is \textit{pseudo-Cauchy} if, for all $\epsilon > 0$, there exists an $N$ such that if $n \geq N$, then $|s_{n+1} - s_n| < \epsilon$.


    Decide which one of the following two propositions is actually true. Supply a proof for the valid statement and a counterexample for the other. \er{
        \item Pseudo-Cauchy sequences are bounded.
        \item If $(x_n)$ and $(y_n)$ are pseudo-Cauchy, then $(x_n + y_n)$ is pseudo-Cauchy as well.
    }
    \sol{\er{
        \item False. Let $s_n = \sum_{i=1}^{n} 1/i$. For any $\epsilon>0$, we can choose $N > 1/\epsilon$, such that when $n> N$, $|s_{n+1} - s_n| = 1/(n+1) < \epsilon$. However, $(s_n) \rightarrow \infty$.
        \item For $\epsilon > 0$, if $n > N$, then \[
        |(x_{n+1} + y_{n+1}) - (x_n + y_n)| = |(x_{n+1} - x_n) + (y_{n+1} - y_n)| \leq |x_{n+1} - x_n| + |y_{n+1} - y_n| < \epsilon/2 + \epsilon/2 = \epsilon\]
    }}
}