\section{The Axiom of Completeness}

\bx{ % Exercise 1.3.1
    \ea{
        \item Write a formal definition in the style of Definition 1.3.2 for the \textit{infimum} or \textit{greatest lower bound} of a set.
        \item Now, state and prove a version of Lemma 1.3.8 for greatest lower bound.
    }
    \sol{
        \ea{
            \item A real number $i$ is the \textit{greatest upper bound} for a set $A \subseteq \mathbb{R}$ if it meets the following two criteria: \er{
                \item $i$ is a lower bound of A;
                \item if $b$ is any lower bound for A, then $i \geq b$.
            }
            \item Assume $i \in \mathbb{R}$ is a lower bound for a set $A \subseteq \mathbb{R}$. Then $i = \inf A$ \textit{if and only if}, for every choice of $\epsilon > 0$, there $\exists a \in A$ satisfying $i + \epsilon > a$.
            \begin{proof}
                Rephrasing the lemma gives us: Given that $i$ is a lower bound, $i$ is the greatest lower bound if and only if any number greater than $i$ is not a lower bound. \er{
                    \item Suppose $i = \inf A$ and consider $i+\epsilon$ for an arbituarily chosen $\epsilon > 0$. Since $i + \epsilon > i$, part (ii) of the definition impliers that $i + \epsilon$ is not a lower bound for $A$. If this is the case, then there must be some element $a \in A$ such that $i + \epsilon > a$.
                    \item Conversely, assume $i$ is a lower bound with the property that for any $\epsilon > 0$, $s+\epsilon$ is not a lower bound of A. Note that this implies that if $b$ is any number more than $i$, then $b$ is not an upperbound. Since we have argued that any larger number than $i$ cannot be a lower bound, if $b$ is some other upper bound for $A$, then $i \geq b$.
                }
            \end{proof}
        }
    }
}

\bx{ % Exercise 1.3.2
    Give an example of each of the following, or state that the request is impossible. \ea{
        \item A set $B$ with $\inf B \geq \sup B$.
        \item A finite set that contains its infimum but not its supremum.
        \item A bounded subset of $\mathbb{Q}$ that contains its supremum but not its infimum.
    }
    \sol{
        \ea{
            \item Possible. Consider the set ${0}$, where $\inf \{0\} = \sup \{0\} = 0$.
            \item Not possible as all finite sets must contain its supremum and infimum.
            \item Possible. Consider $A = \{r \in \mathbb{Q} \hspace{1ex} | \hspace{1ex}1<r\leq2\}$.
        }
    }
}

\bx{ % Exercise 1.3.3
    \ea{
        \item Let $A$ be nonempty and bounded below, and define $B=\{b\in \mathbb{R}: b \text{ is a lower bound for } A\}$. Show that $\sup B = \inf A$.
        \item Use (a) to explain why there is no need to assert that greatest lower bounds exist as part of the Axiom of Completeness.
    }
    \sol{
        \ea{
            \item By definition, $\sup B$ is the greatest lower bound for $A$, meaning it equals $\inf A$.
            \item Part (a) proves the greatest lower bound exists using the least upper bound.
        }
    }
}

\bx{ %Exercise 1.3.4
    Let $A_1, A_2, A_3, \dots$ be a collection of nonempty sets, each of which is bounded above.
    \ea{
        \item Find a formula for $\sup(A_1\cup A_2)$. Extend this to $\sup (\bigcup^n_{k=1}A_k)$.
        \item Consider $\sup (\bigcup^\infty_{k=1} A_k)$. Does the formula in (a) extend to the infinite case?
    }
    \sol{
        \ea{
            \item $\sup (A_1 \cup A_2) = \sup\{\sup A_1, \sup A_2\}$ and $\sup(\bigcup^n_{k=1}A_k) = \sup\{\sup A_k \hspace{1ex} | \hspace{1ex} k = 1, \dots, n \}$
            \item This formula does not extend to infinity. Consider $A_k = [k, k+1]$, where $\bigcup^\infty_{k=1}A_k$ is unbounded.
        }
    }
}

\bx{ % Exercise 1.3.5
    As in Example 1.3.7, let $A\subseteq\mathbb{R}$ be nonempty and bounded above, and let $c \in \mathbb{R}$. This time define the set $cA = \{ca: a \in A\}$.
    \ea{
        \item If $c \geq 0$, show that $\sup(cA) = c \sup A$.
        \item Prostulate a similar type of statement for $\sup(cA)$ for the case $c<0$.
    }

    \sol{
        \ea{
            \item The case of $c=0$ is trivial as it implies that $cA = \{0\}$. Hence $\sup(cA)=c\sup A = 0$.
            
            For $c > 0$, we need to show that $c \sup A$ is the lowest upper bound. Assume $c > 0$. Let $s = c \sup A$. Suppose $ca > s$, then $a > \sup A$ which is impossible, meaning that $s$ is an upper bound on $cA$. Now suppose $s'$ is an upper bound on $cA$ and $s' < s$. Then $s'/c < s/c = \sup A$, meaning $s'/c$ cannot bound $A$. Hence there $\exists a \in A$ such that $s'/c < a$, meaning $s' > ca$, thus $s'$ cannot be an upper bound on $cA$, so $s = c \sup A$ is the least upper bound.
            \item $\sup (cA) = c \inf A$ for $ c < 0$
        }
    }
}

\bx{ % Exercise 1.3.6
    Given sets $A$ and $B$, define $A + B = \{ a+ b: a\in A \text{ and } b \in B\}$. Follow these steps to prove that if $A$ and $B$ are nonempty and bounded above then $\sup (A+B) = \sup A + \sup B$.
    \ea{
        \item Let $s=\sup A$ and $t=\sup B$. Show $s+t$ is an upper bound for $A+B$.
        \item Now let $u$ be an arbituary upper bound for $A+B$, and temporarily fix $a \in A$. Show $t \leq u+a$.
        \item Finally, show $\sup (A+B) = s+t$.
        \item Construct another proof of this same fact using Lemma 1.3.8.
    }

    \sol{
        \ea{
            \item By definition of supremum, $a \leq s$ and $b \leq t$. Adding both equations give $a + b \leq s + t$, hence $s+t$ is an upper bound.
            \item Since $a+b \leq u$ implies $b \leq u-a$, $u-a$ is an upper bound on b, meaning it is greater or equal to the least upper bound of $t$, giving $t \leq u-a$.
            \item From (a), we have shown that $s+t$ is an upper bound for $A+B$, hence it is sufficent to show that $s+t$ is the least upper bound.
            
            Let $u = \sup (A+B)$, from (b) we have $t \leq u - a$ and $s \leq u - b$. Adding and rearranging gives $a + b \leq 2u - s - t$. Since $2u-s-t$ is an upper bound on $A+B$, it must be greater or equal to the least upper bound, giving $u \leq 2u-s-t$, implying $s+t \leq u$. Since $u$ is the least upper bound, $s+t$ must equal $u$.
            \item Showing $s+t-\epsilon$ is not an upper bound for any $\epsilon>0$ proves that it is the least upper bound by Lemma 1.3.8. Rearranging gives $(s-\epsilon/2)+(t-\epsilon/2)$ we know $\exists a > (s-\epsilon/2)$ and $b > (t-\epsilon/2)$, therefore $a+b > s+t-\epsilon$, meaning $s+t$ cannot be made smaller and thus is the least upper bound.
        }
    }
}

\bx{ % Exercise 1.3.7
    Prove that if $a$ is an upper bound for $A$, and if $a$ is also an element of $A$, then it must be that $a = \sup A$.
    \sol{
        Since it is given that $a$ is an upper bound for $A$, we just have to show that $a$ is the least upper bound, meaning any number lower than $a$ would have an $a' \in A$ such that $a' > a$.
        \medskip

        \noindent Suppose $a-\epsilon$ is also an upper bound for $A$ for some $\epsilon > 0$. This is not possible has $a > a'$ and $a \in  A$. Hence by contradiction, $a$ is the lowest upper bound, meaning $a = \sup A$.
    }
}

\bx{ % Exercise 1.3.8
    Compute, without proofs, the suprema and inifima (if they exists) of the following sets: \ea{
        \item $\{m/m:m,n\in \mathbb{N} \text{ with } m<n\}$.
        \item $\{(-1)^m/n:m,n \in \mathbb{N}\}$.
        \item $\{n/(3n+1):n\in\mathbb{N}\}$.
        \item $\{m/(m+n):m,n\in\mathbb{N}\}$.
    }
    \sol{
        \ea{
            \item $\inf = 0$ and $\sup = 1$
            \item $\inf = -1$ and $\sup = 1$
            \item $\inf = 1/4$ and $\sup = 1/3$
            \item $\inf = 0$ and $\sup = 1$
        }
    }
}

\bx{ % Exercise 1.3.9
    \ea{
        \item If $\sup A < \sup B$, show there exists an element that is an upper bound for A.
        \item Give an example to show that this is not always the case if we only assume $\sup A \leq sup B$.
    }
    \sol{
        \ea{
            \item By Lemma 1.3.8, we know there exists a $b$ such that $(\sup B)-\epsilon<b$ for any $\epsilon>0$. We can set $\epsilon$ to be small enough such that $\sup B - \sup A < \epsilon$, implying $\sup A < \sup B - \epsilon < b$ for some $b \in B$, thus $b$ is an upper bound of $A$.
            \item Consider the sets $A = (-\infty, 1]$ and $B = (-\infty, 1)$. No $b \in  B$ is an upperbound since $1 \in A$ and $1 > b$.
        }
    }
}

\bx{ % Exercise 1.3.10
    \textbf{(Cut Property)} The \textit{Cut Property} of the real numbers is the following:

    If $A$ and $B$ are nonempty, disjoint sets with $A \cup B =\mathbb{R}$ and $a<b$ for all $a\in A$ and $b \in B$, then there exists a $c \in \mathbb{R}$ such that $x \leq c$ whenever $x \in A$ and $x \geq c$ whenever $x \in B$.
    \ea{
        \item Use the Axiom of Completeness to prove the Cut Property.
        \item Show that the implication goes the other way; that is, assume $\mathbb{R}$ possesses the Cut Property and let $E$ be a nonempty set that is bounded above. Prove $\sup E$ exists.
        \item The punchline of parts (a) and (b) is that the Cut Property could be used in place of the Axiom of Completeness as the fundamental axiom that distinguishes the real numbers from the rational numbers. To drive this point home, give a concrete example showing that the Cut Property is not a valid statement when $\mathbb{R}$ is replaced by $\mathbb{Q}$.
    }
    \sol{
        \ea{
            \item If $c = \sup A = \inf B$, it is obvious that $a \leq c \leq b$ and we are done. Hence we need to show, by contradiction, that $\sup A < \inf B$ and $\sup A > \inf B$ is false. \er{
                \item \AFSOC $\sup A < \inf B$. We can choose $c = \frac{\sup A + \inf B}{2}$, which satisfies $\sup A < c < \inf B$. Hence it is obvious that $c \notin A$ and $c \notin B$, so $c \notin A \cup B \neq \mathbb{R}$ which is a contradiction.
                \item \AFSOC $\sup A > \inf B$. We can find $a$ such that $a > b$ by subtracting $\epsilon > 0$ and using the definition of supremum and infimum similar to Lemma 1.3.8. Thus creating a contradiction.
            }
            Since both alternatives are impossible, $\sup A = \inf B$.
            \item If $E$ is finite or has a maximum element, that is $\sup E$ and we are done.
            
            Consider the case where $E$ has not maximum element (for example, $\{-1/n : n \in \mathbb{N}\}$). Let $B$ be the sets of all upper bounds of $E$ and let $A = B^c$. It can be said $E \cap B = \emptyset$ otherwise $E$ has a maximum element. Thus $E \subseteq A$.
            
            By the Cut Property, there exists $c$ such that $a \leq c \leq b$ for all $a\in A$ and $b\in B$. Since $c$ is an upper bound on $A$ and $E \subseteq A$, $c$ is also an upper bound on $E$. And since $c \leq b$, $c$ is the lowest upper bound. Therefore, $c = \sup E$.
            \item Consider $A = \{r\in \mathbb{Q} \hspace{1ex}|\hspace{1ex}r^2<2 \text{ or }r < 0\}$, $B=A^c$ does not satisfy the cut property in $\mathbb{Q}$ as $\sqrt{2}\notin \mathbb{Q}$.
        }
    }
}

\bx{ % Exercise 1.3.11
    Decide if the following statements about suprema and infima are true or false. Give a short proof for those that are true. For any that are false, supply an example where the claim in question does not appear to hold. \ea{
        \item If $A$ and $B$ are nonempty, bounded and satisfy $A \subseteq B$, then $\sup A \leq \sup B$.
        \item If $\sup A < \inf B$ for sets $A$ and $B$, then there exists a $c \in \mathbb{R}$ satisfying $a < c <b$ for all $a \in A$ and $b \in B$.
        \item If there exists a $c \in \mathbb{R}$ satisfying $a<c<b$ for all $a \in A$ and $b \in B$, then $\sup A < \inf B$.
    }
    \sol{
        \ea{
            \item True. We know $a \leq \sup A$, $a \leq \sup B$ since $A \subseteq B$. Since $\sup A$ is the least upper bound on A, we have $\sup A \leq \sup B$.
            \item True. Let $c = \frac{\sup A + \inf B}{2}$, $c>\sup A$ implies $a<c$ and $c<\inf B$ implies $c<b$, giving $a<c<b$.
            \item False. Consider $A = (-\infty, 1)$ and $B=(1, \infty)$, $a<1<b$ but $\sup A \nless \inf B$ as $\sup A = \inf B = 1$.
        }
    }
}